Why do we Eval?\\\Large A large-scale study of Eval usage in R

### 2021-03-24 analysis/normalized.Rmd
Function definition and higher-order functions

```{r}
fundefs <- E_cano %>% filter(has_fundef)

nb_fundef_callsites <- fundefs %>% pull(src_ref) %>% n_distinct()
```

There are `r nb_fundef_callsites` call sites with function definitions, i.e. `r ratio(nb_fundef_callsites, nb_call_sites)` %.

@@

r("nb slot access", nb_slot_sites)
r("slot access percent", ratio(nb_slot_sites, nb_call_sites))
r("generalized slot access", nb_generalized_slot_sites)
r("generalized slot access percent", ratio(nb_generalized_slot_sites, nb_call_sites))
```

>>>>

```{r}

fundefs <- E_cano %>% filter(minimized == "FUN")
nb_fundef_sites <- fundefs %>% pull(src_ref) %>% n_distinct()
ndefault_envir_fundefs <- fundefs %>% filter(!is.na(envir))
nb_ndefault_envir_fundef_sites <- ndefault_envir_fundefs %>% pull(src_ref) %>% n_distinct()

generalized_fundefs <- E_cano %>% filter(has_fundef)
nb_generalized_fundef_sites <- generalized_fundefs %>% pull(src_ref) %>% n_distinct()
ndefault_envir_gen_fundefs <- generalized_fundefs %>% filter(!is.na(envir))
nb_ndefault_envir_gen_fundef_sites <- ndefault_envir_fundefs %>% pull(src_ref) %>% n_distinct()

r("function definition sites", nb_fundef_sites)
r("function definition sites percent", ratio(nb_fundef_sites, nb_call_sites))
r("fundef non default envir sites percent", ratio(nb_ndefault_envir_fundef_sites, nb_fundef_sites))
r("generalized function definition sites", nb_generalized_fundef_sites)
r("generalized function definition sites percent", ratio(nb_generalized_fundef_sites, nb_call_sites))
r("generalized funded non default envir sites percent", ratio(nb_ndefault_envir_gen_fundef_sites, nb_generalized_fundef_sites))
```

There are `r nb_generalized_fundef_sites` call sites with function definitions, i.e. `r ratio(nb_generalized_fundef_sites, nb_call_sites)` %.

@@

ndefault_envir_slot <- generalized_slot_access %>% filter(!is.na(envir))
nb_ndefault_envir_slot_sites <- ndefault_envir_slot %>% pull(src_ref) %>% n_distinct()

r("nb slot access", nb_slot_sites)
r("slot access percent", ratio(nb_slot_sites, nb_call_sites))
r("generalized slot access", nb_generalized_slot_sites)
r("generalized slot access percent", ratio(nb_generalized_slot_sites, nb_call_sites))
r("ndefault envir slot site", nb_ndefault_envir_slot_sites)
r("ndefault envir slot site percent", ratio(nb_ndefault_envir_slot_sites, nb_generalized_slot_sites))
```

### 2021-03-24 paper/main.tex
Function definition and higher-order functions

\begin{table}[h]
	\begin{tabular}{|c|c|c|c|c|c|}
		\hline
		Minimized & \#sites & \%sites & \#packages & \#operations & \%envir \\
		\hline
		\texttt{X} & \packageMinimizedcallsitesa  & \packageMinimizedpropsitesa & \packageMinimizedpackagea & \packageMinimizedoperationsaRnd & \packageMinimizedpercentenvira \\
		\hline
		\texttt{F(F())} & \packageMinimizedcallsitesb  & \packageMinimizedpropsitesb & \packageMinimizedpackageb  & \packageMinimizedoperationsbRnd & \packageMinimizedpercentenvirb \\
		\hline
		\texttt{V} & \packageMinimizedcallsitesc & \packageMinimizedpropsitesc & \packageMinimizedpackagec & \packageMinimizedoperationscRnd & \packageMinimizedpercentenvirc \\
		\hline
		\texttt{F(X)}& \packageMinimizedcallsitesd & \packageMinimizedpropsitesd & \packageMinimizedpackaged & \packageMinimizedoperationsdRnd & \packageMinimizedpercentenvird \\
		\hline
		\texttt{model.frame} & \packageMinimizedcallsitese & \packageMinimizedpropsitese & \packageMinimizedpackagee & \packageMinimizedoperationseRnd & \packageMinimizedpercentenvire \\
		\hline
		\texttt{\$}& \packageMinimizedcallsitesf & \packageMinimizedpropsitesf & \packageMinimizedpackagef & \packageMinimizedoperationsfRnd & \packageMinimizedpercentenvirf \\
		\hline
		\texttt{F()}& \packageMinimizedcallsitesg & \packageMinimizedpropsitesg & \packageMinimizedpackageg & \packageMinimizedoperationsgRnd & \packageMinimizedpercentenvirg \\
		\hline
		\texttt{<-} & \packageMinimizedcallsitesh & \packageMinimizedpropsitesh & \packageMinimizedpackageh & \packageMinimizedoperationshRnd & \packageMinimizedpercentenvirh \\
		\hline
		\texttt{FUN} & \packageMinimizedcallsitesi  & \packageMinimizedpropsitesi & \packageMinimizedpackagei & \packageMinimizedoperationsiRnd & \packageMinimizedpercentenviri \\
		\hline
		\texttt{\{BLOCK\}} & \packageMinimizedcallsitesj & \packageMinimizedpropsitesj & \packageMinimizedpackagej & \packageMinimizedoperationsjRnd & \packageMinimizedpercentenvirj \\
		\hline
		\texttt{F(F(<-))} & \packageMinimizedcallsitesk & \packageMinimizedpropsitesk & \packageMinimizedpackagek & \packageMinimizedoperationskRnd & \packageMinimizedpercentenvirk \\
		\hline
	\end{tabular}
	\caption{Minimized expressions: eleven most frequent in number of sites. \#sites is the number of sites; \%sites the percentage of sites across the total number of sites (\packageNbCallSites sites) with at least this minimized expression; \#packages is the number of packages with at least one of the minimized expression and measures how widespread a minimized expression is. \#operations is the average number of steps performed by the R interpreter when given that minimized expression, and \%envir is the percentage of call sites for which the \c{envir} argument is explicitly passed.} \label{tab:minimizedexpressions}
\end{table}

@@

@@

\paragraph{Function definition.} \c{FUN} represents a function definition. It entails very few steps of the interpreter, as the body of the new function is just stored for later execution.

@@

\paragraph{Accessing a slot.}

>>>>

\begin{table}[h]
	\begin{tabular}{|c|c|c|c|c|c|}
		\hline
		Minimized & \#sites & \%sites & \#packages & \#operations & \%envir \\
		\hline
		\texttt{X} & \packageMinimizedcallsitesa  & \packageMinimizedpropsitesa & \packageMinimizedpackagea & \packageMinimizedoperationsaRnd & \packageMinimizedpercentenvira \\
		\hline
		\texttt{F(F())} & \packageMinimizedcallsitesb  & \packageMinimizedpropsitesb & \packageMinimizedpackageb  & \packageMinimizedoperationsbRnd & \packageMinimizedpercentenvirb \\
		\hline
		\texttt{V} & \packageMinimizedcallsitesc & \packageMinimizedpropsitesc & \packageMinimizedpackagec & \packageMinimizedoperationscRnd & \packageMinimizedpercentenvirc \\
		\hline
		\texttt{F(X)}& \packageMinimizedcallsitesd & \packageMinimizedpropsitesd & \packageMinimizedpackaged & \packageMinimizedoperationsdRnd & \packageMinimizedpercentenvird \\
		\hline
		\texttt{model.frame} & \packageMinimizedcallsitese & \packageMinimizedpropsitese & \packageMinimizedpackagee & \packageMinimizedoperationseRnd & \packageMinimizedpercentenvire \\
		\hline
		\texttt{F()}& \packageMinimizedcallsitesf & \packageMinimizedpropsitesf & \packageMinimizedpackagef & \packageMinimizedoperationsfRnd & \packageMinimizedpercentenvirf \\
		\hline
		\texttt{\$}& \packageMinimizedcallsitesg & \packageMinimizedpropsitesg & \packageMinimizedpackageg & \packageMinimizedoperationsgRnd & \packageMinimizedpercentenvirg \\
		\hline
		\texttt{FUN} & \packageMinimizedcallsitesh & \packageMinimizedpropsitesh & \packageMinimizedpackageh & \packageMinimizedoperationshRnd & \packageMinimizedpercentenvirh \\
		\hline
		\texttt{<-} & \packageMinimizedcallsitesi  & \packageMinimizedpropsitesi & \packageMinimizedpackagei & \packageMinimizedoperationsiRnd & \packageMinimizedpercentenviri \\
		\hline
		\texttt{\{BLOCK\}} & \packageMinimizedcallsitesj & \packageMinimizedpropsitesj & \packageMinimizedpackagej & \packageMinimizedoperationsjRnd & \packageMinimizedpercentenvirj \\
		\hline
		\texttt{F(F(<-))} & \packageMinimizedcallsitesk & \packageMinimizedpropsitesk & \packageMinimizedpackagek & \packageMinimizedoperationskRnd & \packageMinimizedpercentenvirk \\
		\hline
	\end{tabular}
	\caption{Minimized expressions: eleven most frequent in number of sites. \#sites is the number of sites; \%sites the percentage of sites across the total number of sites (\packageNbCallSites sites) with at least this minimized expression; \#packages is the number of packages with at least one of the minimized expression and measures how widespread a minimized expression is. \#operations is the average number of steps performed by the R interpreter when given that minimized expression, and \%envir is the percentage of call sites for which the \c{envir} argument is explicitly passed.} \label{tab:minimizedexpressions}
\end{table}

@@

% for this one, currently uses any expression with $, [] or [[]] but we should rather use a more restrictive way (take only the ones without interesting calls)
Minimized expression \texttt{\$} refers to a slot access as the top call in a simple expression, for instance \texttt{a\$slot + 1}. It amounts for \packageGeneralizedSlotAccess sites, \ie \packageGeneralizedSlotAccessPercent of the sites.

@@

\paragraph{Function definition.} \c{FUN} represents a function definition at the top level of the expression, and happens in \packageFunctionDefinitionSites sites, \ie \packageFunctionDefinitionSitesPercent of the sites. It entails very few steps of the interpreter, as the function is returned by \eval, and the body of the new function is just stored for later execution. If we count also function definition inside more complex expressions, it amounts to \packageGeneralizedFunctionDefinitionSites sites, \ie \packageGeneralizedFunctionDefinitionSitesPercent and suggests the widespread use of higher-order functions in R, such as \c{lapply}, the R version of \c{map}.

@@
